\chapter{Energy Model in ns}
\label{chap:enerymodel}

Energy Model, as implemented in \emph{ns}, is a node attribute. The
energy model represents level of energy in a mobile host. The energy
model in a node has a initial value which is the level of energy the node
has at the beginning of the simulation. This is known as
{\tt initialEnergy\_}.
It also has a given energy usage
for every packet it transmits and receives. These are called
{\tt txPower\_} and {\tt rxPower\_}.
The files where the energy model is defined are ~ns/energymodel[.cc and.h].
Other functions/methods described in this chapter may be found in
~ns/wireless-phy.cc, ~ns/cmu-trace.cc, ~ns/tcl/lib[ns-lib.tcl, 
ns-node.tcl, ns-mobilenode.tcl].


\section{The C++ EnergyModel Class}
\label{sec:c++energymodel}

The basic energy model is very simple and is defined by class EnergyModel
as shown below:

\begin{verbatim}
class EnergyModel : public TclObject {
public:
  EnergyModel(double energy) { energy\_ = energy; }
  inline double energy() { return energy\_; }
  inline void setenergy(double e) {energy\_ = e;}
  virtual void DecrTxEnergy(double txtime, double P\_tx) {
    energy\_ -= (P\_tx * txtime);
  }
  virtual void DecrRcvEnergy(double rcvtime, double P\_rcv) {
    energy\_ -= (P\_rcv * rcvtime);
  }
protected:
  double energy\_;
};   
\end{verbatim}

As seen from the EnergyModel Class definition above, there is only a
single class variable {\tt energy\_} which represents the
level of energy in the node at any given time. The constructor
EnergyModel(energy) requires the initial-energy to be passed along as a 
parameter. The other class methods are
used to decrease the energy level of the node for every packet transmitted
( {\tt DecrTxEnergy(txtime, P\_tx)}) and every packet received (
{\tt DecrRcvEnergy (rcvtime, P\_rcv)}) by
the node. {\tt P\_tx} and {\tt P\_rcv} are the
transmitting and
receiving power (respectively) required by the node's interface or PHY.
At the beginning of simulation, {\tt energy\_} is set to
{\tt initialEnergy\_} which is then
decremented for every transmission and reception of packets at the node.
When the energy level at the node goes down to zero, no more packets can
be received or transmitted by the node. If tracing is turned on, line
{\tt DEBUG: node \<node-id\> dropping pkts due to energy = 0}
is printed in the tracefile.


\section{The OTcl interface}
\label{sec:otclenergymodel}

Since the energy model is a node attribute, it may be defined by the
following node configuration APIs:
\begin{verbatim}
$ns_ node-config -energyModel $energymodel \
                 -rxPower $p_rx \
                 -txPower $p_tx \
                 -initialEnergy $initialenergy
\end{verbatim}
Optional values for above configuration parameters of the energy
model are given in the following table:
\begin{table}[h]
\begin{center}
{\tt
\begin{tabular}{|c||c|c|}\hline
{\bf Attribute} & {\bf optional values} & {\bf default values}\\\hline
\rm -energyModel& \rm "EnergyModel"&  \rm none\\\hline
\rm -rxPower& \rm receiving power in watts (e.g 0.3)& \rm 281.8mW\\\hline
\rm -txPower& \rm transmitting power in watts (e.g 0.4)& \rm 
281.8mW\\\hline
\rm -initialEnergy& \rm energy in joules (e.g 0.1)& \rm 0.0\\\hline
\end{tabular}
}
\end{center}
\end{table}
\clearpage

