\chapter{The Class Simulator}
\label{chap:sim}

The overall simulator is described by a Tcl
\clsref{Simulator}{../ns-2/ns-lib.h}.
It provides a set of interfaces for configuring a simulation
and for choosing the type of event scheduler used to drive
the simulation.
A simulation script generally begins by creating an instance
of this class and calling various methods to create nodes, topologies,
and configure other aspects of the simulation.
A subclass of Simulator called {\tt OldSim} is used to support
\emph{ns}~v1 backward compatibility.

The procedures and functions described in this chapter can be found in
\textasciitilde\emph{ns}/{tcl/lib/ns-lib.tcl}, \textasciitilde\emph{ns}/{scheduler.\{cc,h\}}, and, \textasciitilde\emph{ns}/{heap.h}.

\section{Simulator Initialization}
\label{sec:siminit}

When a new simulation object is created in tcl,
\fcnref{the initialization procedure}{../ns-2/ns-lib.h}{Simulator::init}
performs the following operations:
\begin{itemize}
        \item initialize the packet format (calls {\tt create\_packetformat})
        \item create a scheduler (defaults to a calendar scheduler)
        \item create a ``null agent'' (a discard sink used in various places)
\end{itemize}
The packet format initialization sets up field offsets within packets
used by the entire simulation.  It is described in more detail
in \href{the following chapter on packets}{Chapter}{chap:pformat}.
The scheduler runs the simulation in an event-driven manner and may
be replaced by alternative schedulers which provide somewhat
different semantics (see the following section for more detail).
The null agent is created with the following call:
\begin{verbatim}
        set nullAgent\_ [new Agent/Null]
\end{verbatim}
This agent is generally useful as a sink for dropped packets or
as a destination for packets that are not counted or recorded.

\section{Schedulers and Events}
\label{sec:sched}

The simulator is an event-driven simulator.
There are presently four schedulers available in the simulator, each
of which is implemented using a different data structure:
a simple linked-list, heap, calendar queue (default), and a special
type called ``real-time''.  Each of these are described below.
The scheduler runs by selecting the next earliest event, executing
it to completion, and returning to execute the next event.Unit of time used by scheduler is seconds.
Presently, the simulator is single-threaded, and only one event
in execution at any given time.
If more than one event are scheduled to execute at the same time,
their execution is performed on the first scheduled -- first
dispatched manner.  Simultaneous events are not re-ordered anymore by
schedulers (as it was in earlier versions) and all schedulers should
yeild the same order of dispatching given the same input.
 
No partial execution of events or pre-emption is supported.

An {\em event} generally comprises a ``firing time'' and a handler function.
The actual definition of an event is found in \textasciitilde\emph{ns}/{scheduler.h}:
\begin{verbatim}
        class Event { 
        public: 
                Event* next\_;           /* event list / 
                Handler* handler\_;      /* handler to call when event ready /
                double time\_;           /* time at which event is ready /
                int uid\_;               /* unique ID /
                Event() : time\_(0), uid\_(0) {}
        };   
        /*   
         * The base class for all event handlers.  When an event's scheduled
         * time arrives, it is passed to handle which must consume it.
         * {\ie, if it needs to be freed it, it must be freed by the handler.}
         */  
        class Handler {
         public: 
                virtual void handle(Event* event);
        };   
\end{verbatim}
Two types of objects are derived from the base
\clsref{Event}{../ns-2/scheduler.cc}: packets and ``at-events''.
Packets are described in detail 
\href{in the next chapter}{Chapter}{sec:packetclass}.
An at-event is a tcl procedure execution scheduled to occur at
a particular time. 
This is frequently used in simulation scripts.
A simple example of how it is used is as follows:
\begin{verbatim}
        \ldots
        set ns\_ [new Simulator]
        \$ns\_ use-scheduler Heap
        \$ns\_ at 300.5 "\$self complete\_sim"
        \ldots
\end{verbatim}
This tcl code fragment first creates a simulation object,
then changes the default scheduler implementation to be heap-based
(see below), and finally schedules the function {\tt \$self complete\_sim}
to be executed at time 300.5 (seconds)%
(Note that this particular code fragment expects to be encapsulated 
in an object instance procedure, where the appropriate
reference to {\tt \$self} is correctly defined.).
At-events are implemented as events where the handler is
effectively an execution of the tcl interpreter.

\subsection{The List Scheduler}
\label{sec:listsched}

The list scheduler (\clsref{Scheduler/List}{../ns-2/scheduler.cc})
implements the scheduler using a simple linked-list structure.
The list is kept in time-order (earliest to latest), so event
insertion and deletion require scanning the list to find the
appropriate entry.
Choosing the next event for execution requires trimming the first
entry off the head of the list.
This implementation preserves event execution in a FIFO manner
for simultaneous events.

\subsection{the heap scheduler}
\label{sec:heapsched}

The heap scheduler (\clsref{Scheduler/Heap}{../ns-2/scheduler.cc})
implements the scheduler using a heap structure.
This structure is superior to the list structure for a large number
of events, as insertion and deletion times are in $O(\log n)$
for $n$ events.
This implementation in \emph{ns}~v2 is borrowed from the
MaRS-2.0 simulator \cite{Alae94:Design};
it is believed that MaRS itself borrowed the code
from NetSim \cite{Heyb89:Netsim},
although this lineage has not been completely verified.

\subsection{The Calendar Queue Scheduler}
\label{sec:cqsched}

The calendar queue scheduler
(\clsref{Scheduler/Calendar}{../ns-2/scheduler.cc})
uses a data structure analogous to a one-year desk calendar,
in which events on the same month/day of multiple years can be recorded in
one day.
It is formally described in \cite{Brow88:Calendar}, and informally described
in Jain (p. 410) \cite{Jain91:Art}.
The implementation of Calendar queues in \emph{ns}~v2
was contributed by David Wetherall (presently at MIT/LCS).

The calendar queue scheduler since \emph{ns}~v2.33 is improved by the following
three algorithms:
\begin{itemize}
	\item A heuristic improvement that changes the linear search direction
in enqueue operations. The original implementation searches the events in
a bucket in \emph{chronological order} to find the in-order spot for the event
that is being inserted.
The new implementation searches the bucket in \emph{reverse chronological order}
because the event being inserted is usually later than most of the events that are
already in the bucket.
	\item A new bucket width estimation that uses the average interval of
\emph{dequeued events} as the estimation of bucket width. It is stated in
\cite{Brow88:Calendar} that the optimal bucket width should be the \emph{average inverval of all events in the future}.
The original implementation uses the average interval of \emph{future events currently in the most crowded bucket}
as the estimation. This estimation is unstable because it is very likely
that many future events will be inserted into the bucket after this estimation, significantly changing the
averaged event interval in the bucket. The new implementation uses the observed event interval 
in the past, which will not change, to estimate the event interval in future.
	\item SNOOPy Calendar Queue: a Calendar queue variant that dynamically
tunes the bucket width according to the cost trade-off between enqueue
operation and dequeue operation.
The SNOOPy queue improvement is described in \cite{Tan00SNOOPyQueue}.
In this implementation, there is one tcl parameter {\tt adjust\_new\_width\_interval\_ }
specifying the interval with which the SNOOPy queue should re-calculate the bucket width.
Setting this parameter to 0 turns off the SNOOPy queue algorithm and degrades the scheduler
back to the original Calendar Queue. In general, normal simulation users are
not expected to change this parameter.
\end{itemize}
The details of these improvements are described in \cite{WeiCao06NSLinuxTCP}.

The implementation of these three improvements was contributed by Xiaoliang (David) Wei at Caltech/NetLab.

\subsection{The Real-Time Scheduler}
\label{sec:rtsched}

The real-time scheduler (\clsref{Scheduler/RealTime}{../ns-2/scheduler.cc})
attempts to synchronize the execution of events with real-time.
It is currently implemented as a subclass of the list scheduler.
The real-time capability in ns is still under development, but is used
to introduce an \emph{ns}\ simulated network into a real-world topology
to experiment with easily-configured network topologies, cross-traffic, etc. 
This only works for relatively slow network traffic data rates, as the
simulator must be able to keep pace with the real-world packet arrival
rate, and this synchronization is not presently enforced.


\subsection{Precision of the scheduler clock used in ns}
\label{sec:precision}

Precision of the scheduler clock can be defined as the smallest time-scale 
of the simulator that can be correctly represented. The clock variable for 
ns is represented by a double. As per the IEEE std for floating numbers, 
a double, consisting of 64 bits must allocate the following bits between 
its sign, exponent and mantissa fields.

\begin{verbatim}

sign     exponent      mantissa
 1 bit   11 bits        52 bits
\end{verbatim}

Any floating number can be represented in the form ($X * 2^n$) where 
X is the mantissa and n is the exponent. Thus the precision of timeclock in ns can be defined as ($1/2^(52)$).
As simulation runs for longer times the number of remaining bits to represent 
the time educes thus reducing the accuracy. Given 52 bits we can safely say 
time upto around ($2^(40)$) can be represented with considerable accuracy. 
Anything greater than that might not be very accurate as you have remaining 12 
bits to represent the time change. However ($2^(40)$) is a very large number and 
we donot anticipate any problem regarding precision of time in ns.


\section{Other Methods}
\label{sec:other}

The {\tt Simulator} class provides a number of methods used
to set up the simulation.
They generally fall into three categories:
methods to create and manage the topology 
(which in turn consists of
\href{managing the nodes}{Chapter}{chap:nodes} and
\href{managing the links}{Chapter}{chap:links}),
\href{methods to perform tracing}{Chapter}{chap:trace},
and helper functions to deal with the scheduler.
The following is a list of the non-topology related simulator methods:
\begin{verbatim}
Simulator instproc now {} # return scheduler's notion of current time;
Simulator instproc at args # schedule execution of code at specified time;
Simulator instproc cancel args # cancel event;
Simulator instproc run args # start scheduler;
Simulator instproc halt {} # stop (pause) the scheduler;
Simulator instproc flush-trace {} # flush all trace object write buffers;
Simulator instproc create-trace { type files src dst } # create trace object;
Simulator instproc create\_packetformat # set up the simulator's packet format;

\clearpage

\section{Commands at a glance}
\label{sec:simcommand}
\begin{flushleft}
Synopsis:
{\tt ns \\<otclfile\\> \\<arg\\> \\<arg\\>..}\\
Description:
Basic command to run a simulation script in ns.
The simulator (ns) is invoked  via the ns interpreter, an extension of the
vanilla otclsh command shell. A simulation is defined by a OTcl script
(file). Several examples of OTcl scripts can be found under \emph{ns}/tcl/ex
directory.


The following is a list of simulator commands commonly used in simulation
scripts:

{\tt set ns\\_ [new Simulator]}\\
This command creates an instance of the simulator object.


{\tt set now [\\$ns\\_ now]}\\
The scheduler keeps track of time in a simulation. This returns scheduler's
notion of current time.


{\tt \\$ns\\_ halt}\\
This stops or pauses the scheduler.


{\tt \\$ns\\_ run}\\
This starts the scheduler.


{\tt \\$ns\\_ at \\<time\\> \\<event\\>}\\
This schedules an \<event\> (which is normally a piece of code) to be executed
at the specified \<time\>.
e.g \$ns\_ at \$opt(stop) "puts \"NS EXITING..\" ; \$ns\_ halt"
or, \$ns\_ at 10.0 "\$ftp start"


{\tt \\$ns\\_ cancel \\<event\\>}\\
Cancels the event. In effect, event is removed from scheduler's list of 
ready to run events.


{\tt \\$ns\\_ create-trace \\<type\\> \\<file\\> \\<src\\> \\<dst\\> \\<optional arg: op\\>}\\
This creates a trace-object of type \<type\> between \<src\> and \<dst\> objects
and attaches trace-object to \<file\> for writing trace-outputs. If op is defined
as "nam", this creates nam tracefiles; otherwise if op is not defined, ns
tracefiles are created on default.


{\tt \\$ns\\_  flush-trace}\\
Flushes all trace object write buffers.


{\tt \\$ns\\_ gen-map}\\
This dumps information like nodes, node components, links etc created for a
given simulation. This may be broken for some scenarios (like wireless).


{\tt \\$ns\\_ at-now \\<args\\>}\\
This is in effect like command "\$ns\_ at \$now \$args". Note that this function
may not work because of tcl's string number resolution.


These are additional simulator (internal) helper functions (normally used
for developing/changing the ns core code) :

{\tt \\$ns\\_ use-scheduler \\<type\\>}\\
Used to specify the type of scheduler to be used for simulation. The different
types of scheduler available are List, Calendar, Heap and RealTime. Currently
Calendar is used as default.


{\tt \\$ns\\_ after \\<delay\\> \\<event\\>}\\
Scheduling an \<event\> to be executed after the lapse of time \<delay\>.


{\tt \\$ns\\_ clearMemTrace}\\
Used for memory debugging purposes.


{\tt \\$ns\\_ is-started}\\
This returns true if simulator has started to run and false if not.


{\tt \\$ns\\_ dumpq}\\
Command for dumping events queued in scheduler while scheduler is halted.


{\tt \\$ns\\_ create\\_packetformat}\\
This sets up simulator's packet format.

\end{flushleft}
\end{verbatim}

\endinput
